### 2015-05-15 whoop.tex
fix

To make our analysis scale while maintaining precision, and to support recursion, we use the Houdini algorithm~\cite{flanagan2001houdini} to automatically compute summaries (pre- and post-conditions and loop invariants).  Given a pool of \emph{candidate} pre-conditions, post-conditions, and loop invariants, Houdini repeatedly attempts to verify each procedure.  Initially, the entire candidate pool is considered.  If verification fails due to an incorrect candidate, this candidate is discarded.  The process repeats until a fixpoint is reached.  The (possibly empty) set of remaining candidates has been proven to hold, and can used to summarize calls and loops during further program analysis.

Houdini does not generate the initial pool of candidates: \whoop generates them using a set of heuristics, and passes them to Houdini as a starting point.  The idea is to automatically generate likely program invariants based on syntactic patterns extracted from an inter-procedural pass over the code for an entry point.  We give two examples; for clarity we use notation from the simple shared variable concurrent programming model of \S\ref{sec:symbolicpairwise}.  If we observe syntactically that procedure $f$ of entry point $T$ may write to, but does not read from, shared variable $s$, then when instrumenting $T$ with symbol $i$, we guess $s \in \hasbeenwritten{i}$ and $s \notin\hasbeenread{i}$ as post-conditions for $\instrument{f}{i}$.  These guesses may be incorrect, for instance if the potential write to $s$ turns out to be in dead code, or if a read from $s$ has already been issued on entry to $\instrument{f}{i}$.  Similarly, if syntactic analysis indicates that $f$ may unlock mutex $m$, we guess $m \notin \currentlockset{i}$ as a post-condition for $\instrument{f}{i}$; this guess may be wrong, e.g.\ if the unlock operation is not reachable or if a subsequent lock operation acquires the mutex again.  We stress that guessing incorrect candidate invariants does not compromise the soundness of verification: \whoop is free to speculatively generate candidates that are later deemed to be incorrect, and thus discarded, by Houdini.  The balance we try to strike is to have \whoop generate sufficient candidates to enable precise lockset analysis, without generating so many candidates that the speed of the Houdini algorithm is prohibitively slow.

>>>>

To make our analysis scale while maintaining precision, and to support recursion, we use the Houdini algorithm~\cite{flanagan2001houdini} to automatically compute summaries (pre- and post-conditions and loop invariants).  Given a pool of \emph{candidate} pre-conditions, post-conditions, and loop invariants, Houdini repeatedly attempts to verify each procedure.  Initially, the entire candidate pool is considered.  If verification fails due to an incorrect candidate, this candidate is discarded.  The process repeats until a fixpoint is reached.  The (possibly empty) set of remaining candidates has been proven to hold, and can be used to summarize calls and loops during further program analysis.

Houdini does not generate the initial pool of candidates: \whoop generates them using a set of heuristics, and passes them to Houdini as a starting point.  The idea is to automatically generate likely program invariants based on syntactic patterns extracted from an inter-procedural pass over the code for an entry point.  We give two examples; for clarity we use notation from the simple shared variable concurrent programming model of \S\ref{sec:symbolicpairwise}.  If we observe syntactically that procedure $f$ of entry point $T$ may write to, but does not read from, shared variable $s$, then when instrumenting $T$ with symbol $i$, we guess $s \in \hasbeenwritten{i}$ and $s \notin\hasbeenread{i}$ as post-conditions for $\instrument{f}{i}$.  These guesses may be incorrect, for instance if the potential write to $s$ turns out to be in dead code, or if a read from $s$ has already been issued on entry to $\instrument{f}{i}$.  Similarly, if syntactic analysis indicates that $f$ may unlock mutex $m$, we guess $m \notin \currentlockset{i}$ as a post-condition for $\instrument{f}{i}$; this guess may be wrong, for instance if the unlock operation is not reachable or if a subsequent lock operation acquires the mutex again.  We stress that guessing incorrect candidate invariants does not compromise the soundness of verification: \whoop is free to speculatively generate candidates that are later deemed to be incorrect, and thus discarded, by Houdini.  The balance we try to strike is to have \whoop generate sufficient candidates to enable precise lockset analysis, without generating so many candidates that the speed of the Houdini algorithm is prohibitively slow.
